Gains and losses in rheumatoid arthritis data from inferCNV

### 2024-02-16 infercnv/reports/gains_and_losses.Rmd
updating gains_and_losses

```{r get_hmm}
cnv_regions <- xfun::cache_rds({
  cnv_regions <- 
    list.files(
      params$infercnv_dir,
      pattern = 'HMM_CNV_predictions.HMMi6.leiden.hmm_mode-subclusters.Pnorm_0.5.pred_cnv_regions.dat',
      recursive = T) %>%
    purrr::set_names(., dirname(.)) %>%
    purrr::map(function(file) {
      file.path(params$infercnv_dir, file) %>% 
        readr::read_tsv()
      }) %>%
    dplyr::bind_rows(.id = 'sample') %>%
    # refactor chromosomes
    dplyr::mutate(
      chr = chr %>% {forcats::fct_reorder(., gsub('chr', '', .) %>% as.numeric())},
      direction = dplyr::case_when(
        state < 3 ~ 'loss',
        state > 3 ~ 'dupli',
        TRUE ~ 'neutral'),
      region = gsub('.*region_', '', cnv_name) %>% as.numeric())
}, file = 'cnv_regions', rerun = params$rerun)
```

```{r plot_cnv_regions_pred}
pdf('test.pdf', height = 25)
cnv_regions %>%
  dplyr::filter(direction != 'neutral') %>% 
  ggplot(aes(x = region)) +
  geom_bar(position = 'stack') +
  facet_wrap(~ chr, ncol = 1, scales = 'free') +
  theme_classic() +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  theme(legend.position = 'none')
dev.off()
```

@@

```{r plot_chr_gains, fig.width = 10, fig.height = 5}
pdf('test.pdf')
paste0('chr', 1:22) %>% head(2) %>%
  purrr::map(function(chr) {
    sce %>%
      Nebulosa::plot_density(
        paste('proportion', c('dupli', 'loss'), chr, sep = '_'),
        size = 0.5) 
  })
```

@@

  # test differences
  de_res <-
    glmGamPoi::test_de(
      fit, 
      contrast = 
        cond(chr7_gain = 1, cell_type = 'Stromal cells') -
        cond(chr7_gain = 0, cell_type = 'Stromal cells')
    )

@@

>>>>

```{r get_hmm}
cnv_regions <- xfun::cache_rds({
  cnv_regions <- 
    list.files(
      params$infercnv_dir,
      pattern = 'HMM_CNV_predictions.HMMi6.leiden.hmm_mode-subclusters.Pnorm_0.5.pred_cnv_regions.dat',
      recursive = T) %>%
    purrr::set_names(., dirname(.)) %>%
    purrr::map(function(file) {
      file.path(params$infercnv_dir, file) %>% 
        readr::read_tsv()
      }) %>%
    dplyr::bind_rows(.id = 'sample') %>%
    # refactor chromosomes
    dplyr::mutate(
      chr = chr %>% {forcats::fct_reorder(., gsub('chr', '', .) %>% as.numeric())},
      direction = dplyr::case_when(
        state < 3 ~ 'loss',
        state > 3 ~ 'dupli',
        TRUE ~ 'neutral'),
      ploidy = dplyr::case_when(
        state == 1 ~ '0x',
        state == 2 ~ '0.5x',
        state == 3 ~ '1x',
        state == 4 ~ '1.5x',
        state == 5 ~ '2x',
        state == 6 ~ '>2x',
        TRUE ~ NA
        ) %>% forcats::fct_reorder(state),
      region = gsub('.*region_', '', cnv_name) %>% as.numeric(),
      cell_type = cell_group_name %>% gsub('\\..*', '', .))
  cnv_regions$midpoint <-
    (cnv_regions$start + cnv_regions$end) / 2
  cnv_regions
}, file = 'cnv_regions', rerun = params$rerun)
```

```{r plot_cnv_regions_pred, fig.height = 20, fig.width = 10}
p_dat %>%
  ggplot(aes(x = pos, fill = ploidy, group = ploidy)) +
  geom_density(data = p_dat %>% dplyr::filter(direction == 'dupli'),
               aes(y = ..density..), 
               position = position_stack(), 
               colour = NA) +
  geom_density(data = p_dat %>% dplyr::filter(direction == 'loss'),
               aes(y = -..density..), 
               position = position_stack(reverse = T), 
               colour = NA) +
  ggh4x::facet_grid2(chr ~ cell_type, scales = 'free', independent = 'x') +
  theme_bw() +
  theme(axis.text.x = element_blank()) +
  scale_fill_manual(
    values = c('0x' = '#00008B', '0.5x' = '#b4b0ff', '1' = 'white', 
               '1.5x' = '#ffc2c2', '2x' = '#d64242', '>2x' = '#8B0000')) 
```

@@

```{r plot_chr_gains, fig.width = 10, fig.height = 5}
paste0('chr', 1:22) %>% head(2) %>%
  purrr::map(function(chr) {
    sce %>%
      Nebulosa::plot_density(
        paste('proportion', c('dupli', 'loss'), chr, sep = '_'),
        size = 0.5) %>%
      print()
  })
```

@@

  # test differences
  de_res <-
    glmGamPoi::test_de(
      fit, 
      contrast = 
        cond(chr7_gain = 1, cell_type = 'Stromal cell') -
        cond(chr7_gain = 0, cell_type = 'Stromal cell')
    )

@@

Next, we plot the top differentially expressed genes onto the plot of locally amplified / lost regions within chr7 to see where they fall.

```{r plot_des_on_regions}

```
